---
title: "hw3-wenqi"
output: html_document
---

# HW3-Wenqi

* Using the RMarkdown/knitr/github mechanism, implement the following tasks:
1. Use the prostate cancer data.
2. Use the cor function to reproduce the correlations listed in HTF Table 3.1, page 50.
3. Treat lcavol as the outcome, and use all other variables in the data set as predictors.
4. With the training subset of the prostate data, train a least-squares regression model with all predictors using the lm   function.
5. Use the testing subset to compute the test error (average squared-error loss) using the fitted least-squares regression model.
6. Train a ridge regression model using the glmnet function, and tune the value of lambda (i.e., use guess and check to find  the value of lambda that approximately minimizes the test error).
7. Create a figure that shows the training and test error associated with ridge regression as a function of lambda
8. Create a path diagram of the ridge regression analysis, similar to HTF Figure 3.8


```{r}
library('splines')        ## for 'bs'
library('dplyr')          ## for 'select', 'filter', and others
library('magrittr')       ## for '%<>%' operator
library('glmnet')         ## for 'glmnet'
```

# Problem1 : load the data
```{r}
# problem 1
prostate <- 
  read.table(url(
    'https://web.stanford.edu/~hastie/ElemStatLearn/datasets/prostate.data'))

#subset to training examples
prostate_train <- prostate %>%
  filter(train == TRUE) %>% 
  select(-train)
summary(prostate_train)

prostate_test <- prostate %>%
  filter(train == FALSE) %>% 
  select(-train)
```

# Problem2 : correlations

```{r}
cor(prostate_train)
```

# Problem3&4&5: least-squares regression & test error

```{r}
# least-squares regression
fit <- lm(lcavol ~ ., data=prostate_train)

# test error
L2_loss <- function(y, yhat)
  (y-yhat)^2
error <- function(dat, fit, loss=L2_loss)
  mean(loss(dat$lcavol, predict(fit, newdata=dat)))

# train_error 
error(prostate_train, fit)

# testing error
error(prostate_test, fit)
```

# Problem6 : ridge regression

```{r}
# ridge regression
form  <- lcavol ~  lweight + age + lbph + lcp + pgg45 + lpsa + svi + gleason
x_inp <- model.matrix(form, data=prostate_train)
y_out <- prostate_train$lcavol
fit_ridge <- glmnet(x=x_inp, y=y_out, alpha=0,lambda=seq(1, 0, -0.05))

# functions to compute testing/training error with glmnet
error <- function(dat, fit, lam, form, loss=L2_loss) {
  x_inp <- model.matrix(form, data=dat)
  y_out <- dat$lcavol
  y_hat <- predict(fit, newx=x_inp, s=lam)  ## see predict.elnet
  mean(loss(y_out, y_hat))
}

## train_error at lambda=0
error(prostate_train, fit_ridge, lam=0, form=form)
## testing error at lambda=0
error(prostate_test, fit_ridge, lam=0, form=form)
## train_error at lambda=0.01
error(prostate_train, fit_ridge, lam=0.01, form=form)
## testing error at lambda=0.01
error(prostate_test, fit_ridge, lam=0.01, form=form)
## train_error at lambda=0.03
error(prostate_train, fit_ridge, lam=0.03, form=form)
## testing error at lambda=0.03
error(prostate_test, fit_ridge, lam=0.03, form=form)
## train_error at lambda=0.05
error(prostate_train, fit_ridge, lam=0.05, form=form)
## testing error at lambda=0.05
error(prostate_test, fit_ridge, lam=0.05, form=form)
```



## Problem7 : training and test error

```{r}
# compute training and testing errors as function of lambda
err_train_1 <- sapply(fit_ridge$lambda, function(lam) 
  error(prostate_train, fit_ridge, lam, form))
err_test_1 <- sapply(fit_ridge$lambda, function(lam) 
  error(prostate_test, fit_ridge, lam, form))

## plot test/train error
plot(x=range(fit_ridge$lambda),
     y=range(c(err_train_1, err_test_1)),
     xlim=rev(range(fit_ridge$lambda)),
     type='n',
     xlab=expression(lambda),
     ylab='train/test error')
points(fit_ridge$lambda, err_train_1, pch=19, type='b', col='darkblue')
points(fit_ridge$lambda, err_test_1, pch=19, type='b', col='darkred')
legend('topright', c('train','test'), lty=1, pch=19,
       col=c('darkblue','darkred'), bty='n')

colnames(fit_ridge$beta) <- paste('lam =', fit_ridge$lambda)
#print(fit_ridge$beta %>% as.matrix)
```


## Problem8 : path diagram 

```{r}
plot(x=range(fit_ridge$lambda),
     y=range(as.matrix(fit_ridge$beta)),
     type='n',
     xlab=expression(lambda),
     ylab='Coefficients')
for(i in 1:nrow(fit_ridge$beta)) {
  points(x=fit_ridge$lambda, y=fit_ridge$beta[i,], pch=19, col='#00000055')
  lines(x=fit_ridge$lambda, y=fit_ridge$beta[i,], col='#00000055')
}
text(x=0, y=fit_ridge$beta[,ncol(fit_ridge$beta)], 
     labels=rownames(fit_ridge$beta),
     xpd=NA, pos=4, srt=30)
abline(h=0, lty=3, lwd=2)
```


