---
title: "hw5-wenqi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library('MASS') ## for 'mcycle'
library('manipulate') ## for 'manipulate'
library(dplyr)
library(caret)
```

# Seperate data to subsets

```{r}
mcycle <- mcycle
```


```{r}
mcycle_train<-sample_frac(mcycle, 0.75)
sid<-as.numeric(rownames(mcycle_train)) 
mcycle_test<-mcycle[-sid,]
```

#  Predicting the mean acceleration

* Nadaraya-Watson method with the k-NN kernel function

```{r}
y <- mcycle_train$accel
x <- matrix(mcycle_train$times, length(mcycle_train$times), 1)
```

## K-NN function

```{r}
kernel_k_nearest_neighbors <- function(x, x0, k=1) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
```

## Make predictions using the NW method

```{r}
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}
```

## create a grid of inputs 

```{r}
x_plot <- matrix(seq(min(x),max(x),length.out=100),100,1)
```

## make predictions using NW method at each of grid points

```{r}
y_hat_plot <- nadaraya_watson(y, x, x_plot,
  kernel_k_nearest_neighbors)
```

## plot predictions

```{r}
plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot, y_hat_plot, col="#882255", lwd=2) 
```

# AIC & BIC

## Use validation data

```{r}
y <- mcycle_test$accel
x <- matrix(mcycle_test$times, length(mcycle_test$times), 1)
```


## squared-error loss function

```{r}
loss_squared_error <- function(y, yhat)
  (y - yhat)^2
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))
```

## AIC & BIC 

```{r}
aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d
```

## Plot

```{r}
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}

```



```{r}
#manipulate({
   ## make predictions using NW method at training inputs
 #  y_hat <- nadaraya_watson(y, x, x,
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#  edf <- effective_df(y, x, 
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#   aic_ <- aic(y, y_hat, edf)
 #  bic_ <- bic(y, y_hat, edf)
 #  y_hat_plot <- nadaraya_watson(y, x, x_plot,
 #    kern=kernel_k_nearest_neighbors, k=k_slider)
 #  plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
 #  legend('topright', legend = c(
  #   paste0('eff. df = ', round(edf,1)),
  #   paste0('aic = ', round(aic_, 1)),
  #   paste0('bic = ', round(bic_, 1))),
  #   bty='n')
 #  lines(x_plot, y_hat_plot, col="#882255", lwd=2) 
 #}, k_slider=slider(1, 10, initial=3, step=1))

```

# 5-fold cross-validation using the combined training and validation data


```{r}
set.seed(123) 
mcycle_flds  <- createFolds(mcycle$accel, k=5)
print(mcycle_flds)
sapply(mcycle_flds, length) 
```

## 5 estimates of test error 

```{r}
cvknnreg_mcycle <- function(kNN = 10, flds=mcycle_flds) {
  cverr <- rep(NA, length(flds))
  for(tst_idx in 1:length(flds)) { ## for each fold
    
    ## get training and testing data
    mcycle_trn <- mcycle[-flds[[tst_idx]],]
    mcycle_tst <- mcycle[ flds[[tst_idx]],]
    
    ## fit kNN model to training data
    knn_fit <- knnreg(accel ~ times,
                      k=kNN, data=mcycle_trn)
    
    ## compute test error on testing data
    pre_tst <- predict(knn_fit, mcycle_tst)
    cverr[tst_idx] <- mean((mcycle_tst$accel - pre_tst)^2)
  }
  return(cverr)
}
```


# CV-estimated test error

```{r}
## Compute 5-fold CV for kNN = 1:20
cverrs <- sapply(1:30, cvknnreg_mcycle)
print(cverrs) ## rows are k-folds (1:5), cols are kNN (1:20)
cverrs_mean <- apply(cverrs, 2, mean)
cverrs_sd   <- apply(cverrs, 2, sd)

## Plot the results of 5-fold CV for kNN = 1:20
plot(x=1:30, y=cverrs_mean, 
     ylim=range(cverrs),
     xlab="'k' in kNN", ylab="CV Estimate of Test Error")
segments(x0=1:30, x1=1:30,
         y0=cverrs_mean-cverrs_sd,
         y1=cverrs_mean+cverrs_sd)
best_idx <- which.min(cverrs_mean)
points(x=best_idx, y=cverrs_mean[best_idx], pch=30)
abline(h=cverrs_mean[best_idx] + cverrs_sd[best_idx], lty=3)
```


## Explaination

* we could choose the k=23 since it is the best one under 1sd rule.





